Effective Inclusion of Electronic Polarization to a Negatively Charged Membrane

### 2019-03-12 Manuscript/manuscript.tex
Update text

@@

\begin{figure}[tbp!] 
  \centering 
  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
  \caption{\label{fig:POPS-counterions-dens}
    Number density profiles of \ce{K^{+}}, \ce{Na^{+}} counterions along the membrane normal axis 
    for the negatively charged bilayers composed of only POPS (left) or POPC:POPS (5:1, right) 
    using ECC-lipids/ECC-ions (solid lines) and Lipid17/Dang ions (dashed lines).  
    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
}
\todo{JOE: add labels with the relative surface excess $\Gamma$, which are promised to be present in the text.} \\
\todo{JOE: split this figure into two: separate for POPS (to SI) and PC:PS 5:1 (main text)? SAMULI: Should be POPS in main text, mixture in SI.}
\end{figure}

\begin{figure}[tb!] 
  \centering 
  \includegraphics[width=\figwidth]{../Fig/lipids_chemfig_POPC_POPS.pdf} 
  \caption{ \label{fig:chemstruct_pc_ps} 
            The chemical structure of POPC and POPS 
            including the labeling of the carbon segments
            used in the measurements of order parameteres. 
          }
          \todo{The POPS structure and labeling is already in figure \ref{simVSexpNOions_POPS}.
            The POPC structure and labeling could be added on top of the new version of figure \ref{fig:delta_ordPar_NaCl_PC-PS_mix} with only PC,
          where PC results occur in the first time, and this figure could be removed.}
\end{figure}

@@

First, we monitor the response of PC headgroup order parameters to the increasing amount of PS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
According to the electrometer concept, the head group order parameters of PC lipids increase
with the increasing amount of negatively charged PS lipids in experiments \cite{seelig87,scherer87}.
The trend is observed with both force fields in simulations with \ce{K+} counterions and
in ECC-POPS simulation with \ce{Na+} counterions, but not in Lipid17/Dang ion simulations with \ce{Na+} counterions.
The result can be explained by the higher binding affinity of sodium in the
Lipid17/Dang ion simulations (Fig~\ref{fig:POPS-counterions-dens}), which
compensates the negative charge of PS lipids at the interface.
\begin{figure*}[!tbp] 
  \centering 
  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_nacl.pdf} 
  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_kcl.pdf} 
  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_kcl.pdf} 
  \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix} 
    The head group order parameters of a POPC:POPS (5:1) bilayer as a function of PS content
    with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions from simulations with ECC-lipids and Lipid17 \cite{lipid17-future} 
    and experiments (only \ce{Na^+} counterions, but shown also in bottom plots). \cite{roux90}. 
    The underestimated response of ECC-lipids with \ce{Na^+} counterions 
    is likely due to a still slightly overestimated binding affinity of \ce{Na^+} to the phospholipids,
    which is corroborated by the series with \ce{K^+} counterions (lower affinity than \ce{Na^+}),
    where the response is closer to the experiment, which uses only \ce{Na^+} counterions. 
  }
  \todo{JOE+SAMULI: split this figure into separate plots for PC and PS, 
    and change the PS-plot x-axis to reflect the dilution of POPS by PC (now shows the content of PS). 
    The experimental lines in the bottom plot shall be different to make clear that they are not with K counterions (gray dashed?).} \\
  \todo{Error bar of 0.02 should be shown for the experimental POPS results because they come from different experiments, see caption of figure 8 in NMRlipids IV publication.}
\end{figure*}

Second, we monitor the response of PC headgroup order parameters to the additional
amount of sodium and potassium in POPC:POPS (5:1) mixtures.
In experiments, the addition of K$^+$ ions led to a very modest decrease of the order parameters,
while the decrease was more pronounced with Li$^+$ ions and strongest with divalent Ca$^{2+}$ and Mg$^{2+}$ ions \cite{roux90}.
According to the electrometer concept, the decrease of PC lipid head group order parameters
correlates with the amount of bound cations \cite{akutsu81,altenbach84,seelig87,roux90,catte16,nmrlipids_proj4},
thus the binding affinity of the ions to POPC:POPS (5:1) mixtures increase in order K$^{+}$ $<$ Li$^{+}$  $<$ Mg$^{2+}$  $<$ Ca$^{2+}$.
The response of PC headgroup order parameters to the increasing amount of KCl is close to experiments
in both force fields (Fig. \ref{fig:delta_ordPar_KCl_PCPS}), athought the change of the $\alpha$-carbon
order parameter may be slightly overestimated by the Lipid17/Dang simulations. This could be
explained by the slightly deeper penetration of potassium into the bilayer (Fig. \ref{fig:POPS-counterions-dens}).
Because experimental data with the additional amount of sodium in POPC:POPS (5:1) mixture was not available,
we compared the simulation results to the experimental data with the additional amounts of
lithium and potassium ions, which should give the lower and upper bounds to the sodium binding
affinity (Fig. \ref{fig:delta_ordPar_KCl_PCPS}). The PC headgroup response ot the additional
amount of sodium is clearly overestimated by the Lipid17/Dang simulation, which is in line
the higher binding affinity (Fig. \ref{fig:POPS-counterions-dens}). Also, the response
in ECC-POPS simulations is probably too close to the experimental results for lithium,
indicating slight overbinding also in this simulation. In conclusion, the response of PC
headgroups to added amount of PS and to the added amount of ions suggest that the Lipid17/Dang
simulations sinficantly overestimate the sodium binding affinity to the PS containing lipid bilayers, while
ECC-POPS simulations give results closer to the experiments, but may still slightly overestimate
the binding affinity. The binding affinity of potassium is better described by both force fields.

@@

% This§: 
% Other models do not yield good reponse, but
% ECC-lipid describe the changes well
% => Elect. polar. necessarry.
% 
The interaction with \ce{K^+}, which binds very weakly to both neutral and negatively charged membranes, 
renders a qualitatively different response of the order parameter $S^\beta$ in POPS 
in the mixed negatively charged bilayers compared to the neutral bilayers. 
While the order parameter $S^\beta$ increases for both \ce{Na^+} and \ce{Ca^{2+}},
it decreases in the presence of \ce{K^+}
(\ce{KCl}:  Fig.~\ref{fig:delta_ordPar_KCl_PCPS}, 
 \ce{NaCl}: Fig.~\ref{fig:delta_ordPar_NaCl_PCPS}, and
\ce{CaCl2}: Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
Despite the tremendous effort in the development of classical MD models of lipids, however,
there is no such model, 
which would agree with the experimental trends for these salt solutions \cite{nmrlipids_proj4}. 
In contrast, ECC-lipids with ECC-ions capture the different response 
of the order parameters $S^{\beta}$, $S^{\alpha _1}$ and $S^{\alpha _2}$ in POPS 
to various concentrations of \ce{NaCl} and \ce{KCl} in a good agreement with the experiments
(\ce{KCl}:  Fig.~\ref{fig:delta_ordPar_KCl_PCPS}, 
 \ce{NaCl}: Fig.~\ref{fig:delta_ordPar_NaCl_PCPS}). 
Such a dramatic improvement in the structural parameters 
demonstrates that including electronic polarization in phospholipids is crucial 
even for interactions with soft cations like \ce{K^+}.

@@

% This§: 
% adsorption and affinities of Na and K to PC:PS memb
% from densities and rel.surf.excess
% 
%in  Fig.~\ref{fig:POPS-counterions-dens} (only counterions)
%and Fig.~\ref{fig:nacl-dens_PCPS}        (additional salt concentrations), 
%
The affinities of \ce{Na^+} and \ce{K^+} to phospholipid membranes
can be described by their relative surface excesses with respect to water, $\Gamma ^{w} _{ion}$. 
Such a quantity compares the adsorption of ions to the adsorption of water molecules at an interface 
without the necessity of defining a Gibbs dividing surface. \citep{melcr18, chattorajBOOK}
Interestingly enough, \ce{K^+} maintains negative values of $\Gamma^{w}_{K}$ even for the negatively charged PC:PS~(5:1) bilayer.
This is also evident from the density plots in  Fig.~\ref{fig:nacl-dens_PCPS}
showing that the density profile of \ce{K^+} along the membrane normal 
decays clearly faster towards the membrane centre compared to the density of water. 
In contrast, the density profile of \ce{Na^+} has a shoulder 
at the region of the phosphate groups,
which demonstrates that sodium cations adsorb weakly to the membrane interface compared to the aquaeous solution
by yielding a small positive value of $\Gamma^{w}_{Na}$.

\todo{JOE: Add a discussion about the still slightly overestimated affinity of \ce{Na^+}.}

@@

@@

Here, we demonstrate that the same approach 
significantly improves the description of calcium binding 
also to the negatively charged PC:PS (5:1) bilayers. 
In correspondence with the Ref.~\citenum{nmrlipids_proj4},
we also use the changes of the head group order parameters $S^\alpha$ and $S^\beta$ 
in PC and PS lipids 
to quantify the binding affinity of the calcium cations
and the induced structural changes
(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
To directly connect to the previous works \citep{catte16, nmrlipids_proj4}
and to mark the improvement after treating electronic polarizability, 
we show simulation results from the newly developed ECC-lipids and also from Lipid17 \citep{lipid17-future}. 
While the response of the order parameters of PC to \ce{CaCl2} concentrations
is qualitatively correct in the Lipid14/17 model \cite{lipid17},
it is quantitatively comparable to the experimental trends in ECC-lipids. 
Moreover, the structure of PS head group is fundamentaly improved
after accounting for the electronic polarization 
(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}).

Increasing concentrations of \ce{CaCl2} 
induce a systematic decrease of the order parameters $S^\alpha$ and $S^\beta$ 
in both neutral POPC bilayers \cite{catte16, melcr18}
as well as in the negatively charged PC:PS (5:1) bilayers \cite{nmrlipids_proj4, roux90}. 
The change of the head group order parameters 
are well reproduced only by ECC-lipids in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}.

@@

The changes of the order parameters $\alpha$ and $\beta$ at low \ce{CaCl2} concentrations
have a steep onset at low concentrations in the negatively charged bilayers compared to neutral POPC bilayers. 
However, such a detail is not captured by ECC-lipids
indicating the necessity for further improvements.
The source of this discrepancy may arise 
from the interactions between PC and PS lipids and with \ce{Na^+} counterions,
from the sub-optimal structure of the lipids (Fig.~\ref{simVSexpNOions_POPS},
and also possibly from the approximate mean-field treatment of the electronic polarization.

@@

The binding of calcium cations to the phospholipid bilayer is shown 
in the density profiles of \ce{Ca^{2+}}, \ce{Na^+} counterions 
and also \ce{Cl^-} ions in Fig.~\ref{fig:cacl-dens_PCPS}. 
Clearly, after accounting for the electronic polarizability, 
the population of the cations at the membrane interface is significantly diminished. 
\todo{JOE: The density plot still needs to be changed to show both Lipid17 and ECC-lipids, and to support this argument.}
Still, the dominant contribution to the binding of calcium comes from the phosphate groups. 
Interestingly enough,
although the \ce{COO^-} groups increase the affinity of calcium towards PS compared to PC,
 there is no extra density at the outmost surface of the membrane. 
This can be understood from the point of the structural rearangements of the \ce{COO^-} group, 
which take place during the interactions with calcium cations
described above and in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}.

>>>>

Here, we demonstrate that the same approach 
significantly improves the description of calcium binding 
also to the negatively charged PC:PS (5:1) bilayers. 
In correspondence with the Ref.~\citenum{nmrlipids_proj4},
we also use the changes of the head group order parameters $S^\alpha$ and $S^\beta$ 
in PC and PS lipids 
to quantify the binding affinity of the calcium cations
and the induced structural changes
(Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
To directly connect to the previous works \citep{catte16, nmrlipids_proj4}
and to mark the improvement after treating electronic polarizability, 
we show simulation results from the newly developed ECC-lipids and also from Lipid17 \citep{lipid17-future}.

@@

\begin{figure}[tbp!] 
  \centering 
  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_purePOPS_ecclipids-lipid17.pdf}
  \includegraphics[height=\figheight]{../img/ecc_pops/density_profiles_na-k-counterions_wat_phos_compar_5PC-1PS_ecclipids-lipid17.pdf}
  \caption{\label{fig:POPS-counterions-dens}
    Number density profiles of \ce{K^{+}}, \ce{Na^{+}} counterions along the membrane normal axis 
    for the negatively charged bilayers composed of only POPS (left) or POPC:POPS (5:1, right) 
    using ECC-lipids/ECC-ions (solid lines) and Lipid17/Dang ions (dashed lines).  
    The density profiles of phosphate groups and water are divided by 4 and 100, respectively.  
}
\todo{JOE: add labels with the relative surface excess $\Gamma$, which are promised to be present in the text.} \\
\todo{JOE: split this figure into two: separate for POPS (to SI) and PC:PS 5:1 (main text)?} \\
\todo{SAMULI: I feel now that we should keep the as it is, but if something is moved to SI, it should be POPS in main text, mixture in SI.}
\end{figure}

\begin{figure}[tb!] 
  \centering 
  \includegraphics[width=\figwidth]{../Fig/lipids_chemfig_POPC_POPS.pdf} 
  \caption{ \label{fig:chemstruct_pc_ps} 
            The chemical structure of POPC and POPS 
            including the labeling of the carbon segments
            used in the measurements of order parameteres. 
          }
          \todo{The POPS structure and labeling is already in figure \ref{simVSexpNOions_POPS}.
            The POPC structure and labeling could be added on top of the new version of figure \ref{fig:delta_ordPar_NaCl_PC-PS_mix} with only PC,
            where PC results occur in the first time, and this figure could be removed.} \\
          \todo{$\gamma$ carbon of POPS is the COO- minus carbon. The $\gamma$ label should be moved there.}
\end{figure}

@@

First, we monitor the response of POPC headgroup order parameters to the increasing amount of POPS (Fig.~\ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
According to the electrometer concept, the head group order parameters of PC lipids increase
with the increasing amount of negatively charged PS lipids in experiments, because the headgroup P--N dipole
tilts more parallel to the membrane interface \cite{seelig87,scherer87}.
The trend is observed with both force fields in simulations with \ce{K+} counterions and
in ECC-POPS simulation with \ce{Na+} counterions, but not in Lipid17/Dang ion simulations with \ce{Na+} counterions,
where the order parameters are unchanged or decrease and the P--N vector tilts more perpendicular to the membrane interface.
The result can be explained by the higher binding affinity of sodium in the
Lipid17/Dang ion simulations (Fig~\ref{fig:POPS-counterions-dens}), which
compensates the negative charge of PS lipids at the interface.
\begin{figure*}[!tbp] 
  \centering 
  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_nacl.pdf} 
  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_nacl.pdf} 
  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPC_kcl.pdf} 
  \includegraphics[height=\figheightsmall]{../Fig/order_parameters_changes_A-B_PC-PS_mix_POPS_kcl.pdf} 
  \caption{\label{fig:delta_ordPar_NaCl_PC-PS_mix} 
    The head group order parameters of a POPC:POPS (5:1) bilayer as a function of PS content
    with \ce{Na^+} (top) and \ce{K^+} (bottom) counterions from simulations with ECC-lipids and Lipid17 \cite{lipid17-future} 
    and experiments (only \ce{Na^+} counterions, but shown also in bottom plots). \cite{roux90}. 
    The underestimated response of ECC-lipids with \ce{Na^+} counterions 
    is likely due to a still slightly overestimated binding affinity of \ce{Na^+} to the phospholipids,
    which is corroborated by the series with \ce{K^+} counterions (lower affinity than \ce{Na^+}),
    where the response is closer to the experiment, which uses only \ce{Na^+} counterions. 
  }
  \todo{JOE+SAMULI: split this figure into separate plots for PC and PS, 
    and change the PS-plot x-axis to reflect the dilution of POPS by PC (now shows the content of PS). 
    The experimental lines in the bottom plot shall be different to make clear that they are not with K counterions (gray dashed?).} \\
  \todo{Error bar of 0.02 should be shown for the experimental POPS results because they come from different experiments, see caption of figure 8 in NMRlipids IV publication.}
\end{figure*}

Second, we monitor the response of PC headgroup order parameters to the additional
amount of sodium and potassium in POPC:POPS (5:1) mixtures.
In experiments, the addition of K$^+$ ions led to a very modest decrease of the order parameters,
while the decrease was more pronounced with Li$^+$ ions and strongest with divalent Ca$^{2+}$ and Mg$^{2+}$ ions \cite{roux90}.
According to the electrometer concept, the decrease of PC lipid head group order parameters
correlates with the amount of bound cations \cite{akutsu81,altenbach84,seelig87,roux90,catte16,nmrlipids_proj4},
thus the binding affinity of the ions to POPC:POPS (5:1) mixtures increase in order K$^{+}$ $<$ Li$^{+}$  $<$ Mg$^{2+}$  $<$ Ca$^{2+}$.
The response of PC headgroup order parameters to the increasing amount of KCl is close to experiments
and the change in P--N vector is negligible in both force fields (Fig. \ref{fig:delta_ordPar_KCl_PCPS}), athought the change of the $\alpha$-carbon
order parameter may be slightly overestimated by the Lipid17/Dang simulations. This could be
explained by the slightly deeper penetration of potassium into the bilayer (Fig. \ref{fig:POPS-counterions-dens}).
Because experimental data with the additional amount of sodium in POPC:POPS (5:1) mixture was not available,
we compared the simulation results to the experimental data with the additional amounts of
lithium and potassium ions, which should give the lower and upper bounds to the sodium binding
affinity (Fig. \ref{fig:delta_ordPar_NaCl_PCPS}). The PC headgroup response to the additional
amount of sodium is clearly overestimated by the Lipid17/Dang simulation and
the response in ECC-POPS simulations is probably too close to the experimental results for lithium.
In conclusion, the results suggest that the sodium binding affinity (Fig. \ref{fig:POPS-counterions-dens}) and
the tilting of P--N vector more perpendicular to the membrane surface (Fig. \ref{fig:delta_ordPar_NaCl_PCPS})
are clearly overestimated in Lipid17/Dang simulations and slightly overestimated in ECC-POPS simulations,
while 
%In conclusion, the response of PC
%headgroups to added amount of PS and to the added amount of ions suggest that the Lipid17/Dang
%simulations sinficantly overestimate the sodium binding affinity to the PS containing lipid bilayers, while
%ECC-POPS simulations give results closer to the experiments, but may still slightly overestimate
%the binding affinity.
the binding affinity of potassium is better described by both force fields.

@@

% This§: 
% Other models do not yield good reponse, but
% ECC-lipid describe the changes well
% => Elect. polar. necessarry.
% 
The changes of POPS headgroup order parameters are almost negligible in experiments with increasing POPC fraction in bilayer
or monovalent ion concentration in solvent
(Figs. \ref{fig:delta_ordPar_NaCl_PCPS}, \ref{fig:delta_ordPar_KCl_PCPS} and \ref{fig:delta_ordPar_NaCl_PC-PS_mix}),
while calcium leads to relatively large changes with low concentrations followed by a rapid saturation below 100mM (Fig. \ref{fig:delta_ordPar_CaCl_PCPS}) \cite{scherer87,roux90}.
%The interaction with \ce{K^+}, which binds very weakly to both neutral and negatively charged membranes, 
%renders a qualitatively different response of the order parameter $S^\beta$ in POPS 
%in the mixed negatively charged bilayers compared to the neutral bilayers. 
%While the order parameter $S^\beta$ increases for both \ce{Na^+} and \ce{Ca^{2+}},
%it decreases in the presence of \ce{K^+}
%(\ce{KCl}:  Fig.~\ref{fig:delta_ordPar_KCl_PCPS}, 
% \ce{NaCl}: Fig.~\ref{fig:delta_ordPar_NaCl_PCPS}, and
%\ce{CaCl2}: Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
In Lipid17/Dang simulations, the changes of POPS headgroup order parameters with increasing amount of POPC
are overestimated, which is also related to the changes in the orientations of the P--N and C$_{\beta}$--C$_{\gamma}$
vectors (Fig. \ref{fig:delta_ordPar_NaCl_PC-PS_mix}). The overestimation of the order parameter changes
was also observed in previous simulations with other force fields \cite{nmrlipids_proj4}, but not in
ECC-POPS simulation, where also the changes in the PS headgroup orientation are more modest (Fig. \ref{fig:delta_ordPar_NaCl_PC-PS_mix}).
Also, the POPS headgroup orientation and order parameters are more sensitive to the additional amount
of sodium and potassium in Lipid17/Dang simulations, while dependence is weaker and closer to experiments
in ECC-POPS simulations (Figs. \ref{fig:delta_ordPar_NaCl_PCPS} and \ref{fig:delta_ordPar_KCl_PCPS}).
%In contrast, ECC-lipids with ECC-ions capture the different response 
%of the order parameters $S^{\beta}$, $S^{\alpha _1}$ and $S^{\alpha _2}$ in POPS 
%to various concentrations of \ce{NaCl} and \ce{KCl} in a good agreement with the experiments
%(\ce{KCl}:  Fig.~\ref{fig:delta_ordPar_KCl_PCPS}, 
% \ce{NaCl}: Fig.~\ref{fig:delta_ordPar_NaCl_PCPS}). 
%Such a dramatic improvement in the structural parameters 
%demonstrates that including electronic polarization in phospholipids is crucial 
%even for interactions with soft cations like \ce{K^+}. 
In conclusion, using ECC to POPS lipid headgroup improves its
interactions with monovalent ions and POPC headgroups when compared with NMR order parameter data,
althought there is room for improvement in the sodium binding affinity.

@@

%SAMULI: I WOULD LEAVE THIS PART OUT
%% This§: 
%% adsorption and affinities of Na and K to PC:PS memb
%% from densities and rel.surf.excess
%% 
%%in  Fig.~\ref{fig:POPS-counterions-dens} (only counterions)
%%and Fig.~\ref{fig:nacl-dens_PCPS}        (additional salt concentrations), 
%%
%The affinities of \ce{Na^+} and \ce{K^+} to phospholipid membranes
%can be described by their relative surface excesses with respect to water, $\Gamma ^{w} _{ion}$. 
%Such a quantity compares the adsorption of ions to the adsorption of water molecules at an interface 
%without the necessity of defining a Gibbs dividing surface. \citep{melcr18, chattorajBOOK}
%Interestingly enough, \ce{K^+} maintains negative values of $\Gamma^{w}_{K}$ even for the negatively charged PC:PS~(5:1) bilayer.
%This is also evident from the density plots in  Fig.~\ref{fig:nacl-dens_PCPS}
%showing that the density profile of \ce{K^+} along the membrane normal 
%decays clearly faster towards the membrane centre compared to the density of water. 
%In contrast, the density profile of \ce{Na^+} has a shoulder 
%at the region of the phosphate groups,
%which demonstrates that sodium cations adsorb weakly to the membrane interface compared to the aquaeous solution
%by yielding a small positive value of $\Gamma^{w}_{Na}$.

@@

The headgroup order parameters of PC lipids decrease more upon addition of \ce{CaCl2}
in POPC:POPS (5:1) mixtures than in pure PC bilayers due to the higher calcium binding
affinity to the negatively charged lipid bilayers \cite{akutsu81,altenbach84,roux90,nmrlipids_proj4}.
Following our previous work \cite{nmrlipids_proj4}, we use the experimental
headgroup order parameter data as a function of CaCl$_2$ concentration
from POPC:POPS (5:1) mixture to evaluate the calcium binding affinity to lipid bilayers containing PS lipids.
The decrease of the PC headgroup order parameters upon addition of \ce{CaCl2} 
is overestimated in the Lipid17/Dang simulation, while ECC-POPS simulations give results
closer to the experiments (Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}). 
This indicates that the lower binding affinity of calcium in ECC-POPS
simulation is more realistic (Fig.~\ref{fig:cacl-dens_PCPS}).
\todo{JOE: The density plot still needs to be changed to show both Lipid17 and ECC-lipids, and to support this argument.}
However, the steep decrease of the order parameters at low calcium concentrations
observed especially in the negatively charged bilayers is not captured by the ECC-lipid simulations.

%The binding of calcium cations to the phospholipid bilayer is shown 
%in the density profiles of \ce{Ca^{2+}}, \ce{Na^+} counterions 
%and also \ce{Cl^-} ions in Fig.~\ref{fig:cacl-dens_PCPS}. 
%Clearly, after accounting for the electronic polarizability, 
%the population of the cations at the membrane interface is significantly diminished.

@@

@@

The source of this discrepancy may arise from the interactions between PC and PS lipids and with \ce{Na^+} counterions,
from the sub-optimal structure of the lipids (Fig.~\ref{simVSexpNOions_POPS},
and also possibly from the approximate mean-field treatment of the electronic polarization.

@@

Still, the dominant contribution to the binding of calcium comes from the phosphate groups. 
Interestingly enough,
although the \ce{COO^-} groups increase the affinity of calcium towards PS compared to PC,
 there is no extra density at the outmost surface of the membrane. 
This can be understood from the point of the structural rearangements of the \ce{COO^-} group, 
which take place during the interactions with calcium cations
described above and in Fig.~\ref{fig:delta_ordPar_CaCl_PCPS}.
